Concurrent Programming with Harmony

### 2020-07-29 book/paper.tex
Latest

It is instructive to see what happens when a writer is in the critical
section and two readers try to enter.  The first reader successfully
acquires \textit{rlock} and but blocks when trying to acquire
\textit{rwlock}, which is held by the writer.  The second reader blocks
trying acquire \textit{rlock} because it is held by the first reader.
When the writer leaves the critical section, the first reader acquires
\textit{rwlock}, sets \textit{nreaders} to 1, and releases \textit{rlock}.
\textit{rwlock} is still held.
Then the second reader acquires \textit{rlock} and, assuming the first
reader is still in the critical section, increments \textit{nreaders} to 2
and enters the critical section \emph{without} acquiring \textit{rwlock}.
\textit{rwlock} is essentially jointly held by both readers.
It does not matter in which order they leave: the second will release
\textit{rwlock}.

While Harmony does not detect any issues, we must remember what it is checking
for: it checks to make sure that there are never a reader and a writer
in the critical section, and there are never two readers in the critical
section.  Moreover, it checks progress: all executions eventually terminate.
What it does \emph{not} check is if the code allows more than one reader
in the critical section.  Indeed, we could have implemented the reader/writer
lock as in \autoref{fig:rwmulti} using an ordinary lock, never allowing
more than one process in the critical section.

@@

\section*{Exercises}
\begin{problems}
\item Why do you suppose the code in \autoref{fig:rwmulti} uses a flag
per process rather than a simple counter?
Can you write a version of this test that uses a counter instead
of a flag per process?
% \item Write a ``library implementation'' of reader/writer locks so multiple
% reader/writer locks can be instantiated.
\item Write a test program for ``bound locks'' (Exercise~\ref{ex:boundlock})
that checks that up to \texttt{M} processes
can acquire the bound lock at the same time.
\item It can be useful for a process to \emph{downgrade} its write lock
to a read lock when its done writing but not done reading.  Implement
\texttt{downgrade}.  Note, you cannot do this simply by releasing the
write lock and then obtaining the read lock.
\end{problems}

@@

\section*{Exercises}
\begin{problems}
\item Assume that you have only binary semaphores.
Implement counting semaphores using binary semaphores using split binary semaphores.
\item Implement a solution to the producer/consumer problem
(\autoref{fig:boundedbuffer}) using split binary semaphores.
\item \label{ex:onelane} Cornell's campus features some one-lane bridges.
Cars can only go in one direction at a time. Consider northbound
and southbound cars wanting to cross a one-lane bridge.
The bridge allows arbitrary many cars, as long as they're going in the
same direction.
Implement a lock that observes this requirement using SBS.  Write methods
\texttt{nb\_enter()} that a car must invoke before going northbound on
the bridge and \texttt{nb\_leave()} that the car must invoke after leaving
the bridge.  Similarly write \texttt{sb\_enter()} and \texttt{sb\_leave()}
for southbound cars.

>>>>

It is instructive to see what happens when a writer is in the critical
section and two readers try to enter.  The first reader successfully
acquires \textit{rlock} but blocks when trying to acquire
\textit{rwlock}, which is held by the writer.  The second reader blocks
trying acquire \textit{rlock} because it is held by the first reader.
When the writer leaves the critical section, the first reader acquires
\textit{rwlock}, sets \textit{nreaders} to 1, and releases \textit{rlock}.
\textit{rwlock} is still held.
Then the second reader acquires \textit{rlock} and, assuming the first
reader is still in the critical section, increments \textit{nreaders} to 2
and enters the critical section \emph{without} acquiring \textit{rwlock}.
\textit{rwlock} is essentially jointly held by both readers.
It does not matter in which order they leave: the second will release
\textit{rwlock}.

While Harmony does not detect any issues, we must remember what it is checking
for: it checks to make sure that there are never a reader and a writer
in the critical section, and there are never two writers in the critical
section.  Moreover, it checks progress: all executions eventually terminate.
What it does \emph{not} check is if the code allows more than one reader
in the critical section.  Indeed, we could have implemented the reader/writer
lock using an ordinary lock, never allowing
more than one process in the critical section, and Harmony would not have
complained.

@@

\section*{Exercises}
\begin{problems}
\item Why do you suppose the code in \autoref{fig:rwmulti} uses a flag
per process rather than a simple counter?
Can you write a version of this test that uses a counter instead
of a flag per process?
% \item Write a ``library implementation'' of reader/writer locks so multiple
% reader/writer locks can be instantiated.
\item Write a test program for ``bound locks'' (Exercise~\ref{ex:boundlock})
that checks that up to \texttt{M} processes
can acquire the bound lock at the same time.
\end{problems}

@@

\section*{Exercises}
\begin{problems}
\item Assume that you have only binary semaphores.
Implement counting semaphores using split binary semaphores.
\item With reader/writer locks,
it can be useful for a process to \emph{downgrade} its write lock
to a read lock when its done writing but not done reading.  Implement
a \texttt{downgrade} method.  Note, you cannot do this simply by releasing the
write lock and then obtaining the read lock, because the process needs
to remain in the critical section.
\item Implement a solution to the producer/consumer problem
(\autoref{fig:boundedbuffer}) using split binary semaphores.
\item \label{ex:onelane} Cornell's campus features some one-lane bridges.
Cars can only go in one direction at a time. Consider northbound
and southbound cars wanting to cross a one-lane bridge.
The bridge allows arbitrary many cars, as long as they're going in the
same direction.
Implement a lock that observes this requirement using SBS.  Write methods
\texttt{nb\_enter()} that a car must invoke before going northbound on
the bridge and \texttt{nb\_leave()} that the car must invoke after leaving
the bridge.  Similarly write \texttt{sb\_enter()} and \texttt{sb\_leave()}
for southbound cars.
